---
title: "IBM Attriration-Kaggle"
author: "sidharth Khatua"
date: "30 April 2018"
output:
  html_document: default
---
#Description of the Dataset

Link to the website :- <https://www.kaggle.com/pavansubhasht/ibm-hr-analytics-attrition-dataset>

Uncover the factors that lead to employee attrition and explore important questions such as 'show me a breakdown of distance from home by job role and attrition' or 'compare average monthly income by education and attrition'. This is a fictional data set created by IBM data scientists.

Education 1 'Below College' 2 'College' 3 'Bachelor' 4 'Master' 5 'Doctor'

EnvironmentSatisfaction 1 'Low' 2 'Medium' 3 'High' 4 'Very High'

JobInvolvement 1 'Low' 2 'Medium' 3 'High' 4 'Very High'

JobSatisfaction 1 'Low' 2 'Medium' 3 'High' 4 'Very High'

PerformanceRating 1 'Low' 2 'Good' 3 'Excellent' 4 'Outstanding'

RelationshipSatisfaction 1 'Low' 2 'Medium' 3 'High' 4 'Very High'

WorkLifeBalance 1 'Bad' 2 'Good' 3 'Better' 4 'Best'


##Import library 
```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(caret)
library(rpart)
library(rattle)
library(Amelia)
library(plotly)
library(gridExtra)
library(VIM)
library(ROCR)
library(randomForest)
```
##Import the dataset
```{r}
ibm <- read.csv('IBM.csv',stringsAsFactors = TRUE,na.strings = c(' ','',NA))
dim(ibm)
table(ibm$Attrition)
237/(237+1233)
str(ibm)
```
##Check for missing variables 
we use **missmap()** to visualise the missingness in the data.
```{r}
missmap(ibm,y.at = c(1),y.labels = c(''))
```

Above chart indicates that there is no missingness in the data.

##EDA analysis on the dataset.


```{r,fig.width=10}
a <-ggplot(ibm,aes(ï..Age,fill=Attrition))+
  geom_density(color='black',alpha=.5)+
  ggtitle('Histogram of Age vs Attrition')+
  theme_bw()+
  xlab('Age')
b <- ggplot(ibm,aes(MonthlyIncome,fill=Attrition))+
  geom_density(color='black',alpha=.5)+
  ggtitle('Histogram of MonthlyIncome vs Attrition')+
  theme_bw()+
  xlab('MonthlyIncome')
c <- ggplot(ibm,aes(HourlyRate,fill=Attrition))+
  geom_density(color='black',alpha=.5)+
  ggtitle('Histogram of HourlyRate vs Attrition')+
  theme_bw()+
  xlab('HourlyRate')
d <- ggplot(ibm,aes(DailyRate,fill=Attrition))+
  geom_density(color='black',alpha=.5)+
  ggtitle('Histogram of DailyRate vs Attrition')+
  theme_bw()+
  xlab('DailyRate') 
grid.arrange(a,b,c,d)

```
From the above diagram we can observe that:

1.	Age at **approx 30** have higher attriration rate.
2.	**lower monthly income salary employee** churn higher compared to others.
3.	Emloyee with **low Hourly rate and Daily Rate** churn higher compared to others.

```{r,fig.width=10,message= FALSE}
e <-ggplot(ibm,aes(ibm$JobRole,fill=Attrition))+
  geom_bar(color='black',alpha=.5)+
  ggtitle('JobRole vs Attrition')+
  theme_bw()+
  xlab('jobrole')+
  coord_flip()

f <- ggplot(ibm,aes(ibm$EducationField,fill=Attrition))+
  geom_bar(color='black',alpha=.5)+
  ggtitle('Education Field vs Attrition')+
  theme_bw()+
  xlab('jobrole')+
  coord_flip()

g <-ggplot(ibm,aes(ibm$NumCompaniesWorked,fill=Attrition))+
  geom_histogram(color='black',alpha=.5)+
  ggtitle('No of Companies Worked vs Attrition')+
  theme_bw()+
  xlab('No of Companies Worked')+
  scale_x_continuous(breaks = seq(0,15,1))

h <-ggplot(ibm,aes(ibm$WorkLifeBalance,fill=Attrition))+
  geom_bar(color='black',alpha=.5)+
  ggtitle('Worklifebalance vs Attrition')+
  theme_bw()+
  xlab('Worklifebalance')+
  coord_flip()

i <- ggplot(ibm,aes(ibm$Gender,fill=Attrition))+
  geom_bar(color='black',alpha=.5)+
  ggtitle('Gender & marital status vs Attrition')+
  theme_bw()+
  xlab('Gender')+
  coord_flip()+facet_wrap(~MaritalStatus)
i
grid.arrange(e,f,g,h)


```

From the above Diagram we can observe that:

1.	Single Male churn higher compared to others.
2.	Lab tech, Sales executives, Research scientist churn higher compared to others.
3.	Employee with life sciences and Medical tend to churn higher.
4.	Employee with 1 company worked with before tend to churn higher.
5.	Employee with better work life balance churn high compared to others.

```{r}
ggplot(ibm,aes(ibm$YearsAtCompany,ibm$YearsSinceLastPromotion,color=ibm$OverTime))+
  geom_point()+
  facet_wrap(~ibm$Attrition)
```

From the above Diagram we can observe that:

1.	The longer you stay in the company the longer is the years since last promotion.

##Data spliting.

we will split the data into two set **train** and **test** set of ratio **70:30**.

we will train the model on **train** set and validate the model on **test** set.

```{r}
sample <- createDataPartition(ibm$Attrition,times = 1,p = .7,list = FALSE)
ibm.train <- ibm[sample,]
ibm.test <- ibm[-sample,]
```
##Model Used: [Decision Tree](https://www.analyticsvidhya.com/blog/2016/04/complete-tutorial-tree-based-modeling-scratch-in-python/)

```{r,fig.width=15,fig.height=15}
set.seed(100)
model.rpart <- rpart(Attrition~.,data = ibm.train)
fancyRpartPlot(model.rpart,main = 'Attiration Model',type =2)
predict.y <- predict(model.rpart,newdata = ibm.test[,-2],type = 'class')
#Validation 
confusionMatrix(ibm.test$Attrition,predict.y)

```

1:We have tried **Pruning the model** to simplify the model.

2:Used **plotcp()** to get the graph where we can establish the **CP compelixity parameter.**

```{r}
##pruning the data:
plotcp(model.rpart,minline = TRUE)
model.rpart.prune <- prune(model.rpart,cp = .038)

##Visulalize Prune Descision Tree 
fancyRpartPlot(model.rpart.prune,main = "Attiration Model")
predict.prune <- predict(model.rpart.prune,type = 'class',newdata = ibm.test[,-2])

#Validation using confusion matrix.
confusionMatrix(ibm.test$Attrition,predict.prune)
```

##Model Used: [Random Forest](https://www.analyticsvidhya.com/blog/2014/06/introduction-random-forest-simplified/)

```{r}

set.seed(101)
model.rf <- randomForest(Attrition~.,data = ibm.train)

predict.rf <- predict(model.rf,newdata = ibm.test[,-2])

#validate
confusionMatrix(predict.rf,ibm.test$Attrition)

#Plot the Error Metric 
plot(model.rf)
```

##Important variables related to Atrriration rate.

```{r,fig.width=10,fig.height=10}
varImpPlot(model.rf,main = 'Important Variables in Random Forest.')
```










